PAMELA: an annotation-based Java Modeling Framework

### 2021-02-12 6-Experimentations.tex
sync with overleaf

Related to the security pattern implementation, PAMELA enables the definition of additional security behavior to existing Java code.
Patterns are defined in PAMELA using three classes, each one representing a different conceptual level PatternFactory, PatternDefinition, PatternInstances.

To declare a Pattern on existing code, pattern elements such as Pattern Stakeholders and methods need to be annotated with provided security pattern-specific annotations. These annotations will be discovered at run-time by the PatternFactory and stored in PatternDefinition attributes.

>>>>

Related to the security pattern implementation, PAMELA enables the definition of additional security behavior to existing Java code.
Patterns are defined in PAMELA using three classes, each one representing a different conceptual level \texttt{PatternFactory}, \texttt{PatternDefinition}, \texttt{PatternInstance}.

To declare a Pattern on existing code, pattern elements such as Pattern Stakeholders and methods need to be annotated with provided security pattern-specific annotations. These annotations will be discovered at run-time by the \texttt{PatternFactory} and stored in \texttt{PatternDefinition} attributes.

### 2021-02-12 architecture.tex
sync with overleaf

%\begin{figure}
%    \centering
\begin{lstlisting}[language=Java,basicstyle=\ttfamily\footnotesize, caption=model execution/manipulation, label=lst:execution]
// Instantiate the meta-model
// by computing the closure of concepts graph
ModelContext modelContext 
    = ModelContextLibrary.getModelContext(Library.class);
// Instantiate the factory
ModelFactory factory  = new ModelFactory(modelContext);
// Instantiate a Library
Library myLibrary = factory.newInstance(Library.class);
myLibrary.setName("My library");
// Instantiate some Books
Book myFirstBook 
    = factory.newInstance(Book.class, "Lord of the ring");
Book anOtherBook = factory.newInstance(Book.class, "Holy bible");
myLibrary.addToBooks(myFirstBook);
myLibrary.addToBooks(anOtherBook);
\end{lstlisting}
 %   \caption{Executing a PAMELA model}
 %   \label{fig:ExecutingPamelaModel}
%\end{figure}

>>>>

%\begin{figure}
%    \centering
\begin{lstlisting}[language=Java,basicstyle=\ttfamily\footnotesize, caption=model execution/manipulation, label=lst:execution]
// Instantiate the meta-model
// by computing the closure of concepts graph
ModelContext modelContext 
    = ModelContextLibrary.getModelContext(Library.class);
// Instantiate the factory
ModelFactory factory  = new ModelFactory(modelContext);
// Instantiate a Library
Library myLibrary = factory.newInstance(Library.class);
// Instantiate some Books
Book myFirstBook 
    = factory.newInstance(Book.class, "Lord of the ring");
Book anOtherBook = factory.newInstance(Book.class, "Holy bible");
myLibrary.addToBooks(myFirstBook);
myLibrary.addToBooks(anOtherBook);
\end{lstlisting}
 %   \caption{Executing a PAMELA model}
 %   \label{fig:ExecutingPamelaModel}
%\end{figure}

### 2021-02-12 implementation.tex
sync with overleaf

The PAMELA framework is a 100\% pure Java (\textgreater 1.5), compilable by a classical Java compiler and executable in a classical Java Virtual Machine.

>>>>

The PAMELA framework is a 100\% pure Java ($\geq$ /1.81.8), compilable by a classical Java compiler and executable in a classical Java Virtual Machine.

### 2021-02-12 pamelaProcess.tex
sync with overleaf

We advocate for a strong coupling between model and source-code, to give architects and developers a way to both interact during the whole development cycle. PAMELA is an annotation-based Java modeling framework providing a smooth integration between model and code, without code generation nor externalized model serialization. \sg{Instead of generating the code, the API (mostly Java interfaces with abstract method declarations) is locally executed.}. The idea is to avoid separation between modeling and code to facilitate consistency management and to /avoid round-trip issues. Figure \ref{fig:PamelaVision} summarizes the PAMELA architecture.

@@

\sg{
Here are various conceivable scenarios for PAMELA use:
\begin{itemize}
    \vspace{-0.2cm}\item \textbf{Programming use} : programmers reuse existing annotations and write model and code at the same time (the modelers and programmers share the same artifacts : the Java code). This scenario includes the case where the model (made by the modelers) is pre-existing.
    \vspace{-0.2cm}\item \textbf{Reengineering} : programmers start from an existing base of code (legacy) and refactor it while replacing existing code by abstract method declarations in Java interface, reducing risks of errors.
    \vspace{-0.2cm}\item \textbf{Aspect-oriented programming} : programmers may use or redefine "patterns" (eg Security Patterns) which offers code weaving at runtime, as well as runtime monitoring.
    \vspace{-0.2cm}\item \textbf{Advanced programmers use} : programmers may extends PAMELA with their own annotations, their own implementations, their own patterns, etc...
\end{itemize}
}

>>>>

We advocate for a strong coupling between model and source-code, to give architects and developers a way to both interact during the whole development cycle. PAMELA is an annotation-based Java modeling framework providing a smooth integration between model and code, without code generation nor externalized model serialization. \sg{Instead of generating the code, the API (mostly Java interfaces with abstract method declarations) is locally executed (interpreted).} The idea is to avoid separation between modeling and code to facilitate consistency management and to /avoid round-trip issues. Figure \ref{fig:PamelaVision} summarizes the PAMELA architecture and is more detailed thereafter.

@@

\sg{
Here are various conceivable scenarios for PAMELA use:
\begin{itemize}
    \vspace{-0.2cm}\item \textbf{Programming use} : programmers reuse existing annotations and write model and code at the same time (the modelers and programmers share the same artifacts : the Java code). This scenario includes the case where the model (made by the modelers) is pre-existing.
    \vspace{-0.2cm}\item \textbf{Reengineering} : programmers start from an existing base of code (legacy) and refactor it while replacing existing code by abstract method declarations in Java interface, reducing risks of errors.
    \vspace{-0.2cm}\item \textbf{Aspect-oriented programming} : programmers may use or redefine "patterns" (eg Security Patterns \cite{silva20}) which offers code weaving at runtime, as well as runtime monitoring.
    \vspace{-0.2cm}\item \textbf{Advanced programmers use} : programmers may extends PAMELA with their own annotations, their own implementations, their own patterns, etc...
\end{itemize}
}
